Trending assessment for nowcasts, measurements, and forecasts

### 2024-03-08 sections/03-trending.tex
Update on Overleaf.

Third, for measurement, a test method is usually evaluated against a gold standard to test whether the test method could replace the gold standard.
Thus, the gold standard measurements are not available if the quality of the test method is sufficient. \improvement{irgendwie für mich nicht klar formuliert: man unteruscht ja bei meausurement nur ob die gemessenen Änderungen von den zwei Methoden konsistent sind oder?}
Thus, the predicted change is computed based on the values of the test method at time $t$ and $t + l$.

@@

\subsection{Accounting for noise and non-informative small changes}\label{subsec:trending-noise}

@@

\subsection{Bootstrap confidence intervals and the conditional trending plot}\label{subsec:trending-cond-prob}
\unsure{Vielleicht doch zwei subsections?}
Confidence intervals can account for the estimation uncertainty of the measures above.
Bootstrap confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
New samples are drawn with replacement from the dataset and are not based on parametric assumptions as classical confidence intervals are.
The confidence interval is then computed based on the bootstrap samples.
We examine three methods for bootstrapping: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
In the \textit{percentile} approach, the confidence interval for the level $\alpha$ is built directly from the empirical distribution of the bootstrap estimators.
The \textit{basic} approach computes the confidence interval based on the non-bootstrap estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but requires smaller datasets for reasonable confidence intervals.
The basic approach balances these two objectives.
We compare the approaches in a small synthetic data study on their small-dataset behavior and computation time in Appendix~\ref{sec:appendix-trending}.
\Ac{bca} is the only method that maintains the confidence level also for small datasets while increasing the computation time only moderately for larger datasets.
Therefore, we use the \ac{bca} method for confidence intervals in the applications in Section~\ref{sec:application}.

@@

@@

\begin{figure}
    \centering
    \begin{subfigure}{.48\textwidth}
        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_asym_butterfly}
        \caption{First dataset with asymmetric dependence.}
    \end{subfigure}
    \begin{subfigure}{.48\textwidth}
        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_normal}
        \caption{Second dataset. }
    \end{subfigure}
    \caption{Resulting conditional trending plot for different bandwidth selection processes. Cross-validation least squares takes a considerably larger computation time. It does not converge for the first data set and yields a bandwidth too small for the second data set. The rule of thumb is the fastest method but tends to oversmooth. The cross-validation maximum likelihood method yields a more reasonable bandwidth with moderate computation time. }\label{fig:trending-cond-prob-bw}
\end{figure}

@@

If the true value is given as a distribution because it is still unknown, the probabilities $p_t$ can be computed by integration.
Let for two nowcasts the distributions be given by \acp{pdf} $f_{t+l|t+l}$ and $f_{t|t+l}$ with \acp{cdf}  $F_{t+l|t+l}$ and $F_{t|t+l}$.
Then, the probability of a negative change can be computed by
\begin{align}
    \bar{p}_t
        &= \int_{\begin{subarray}{l}x_1, x_2 \in \R: \\ x_2 < x_1\end{subarray}} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: (x_1, x_2) \nonumber\\
        &= \int_{x_1 \in \R} \int_{-\infty}^{x_1} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \nonumber\\
        &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \label{eq:trending-probabilistic-pdf}.
\end{align} \unsure{vielleicht eher d(x,y) wegen einem Integralzeichen?}

\unsure{Die Unabhängigkeitsannahme ist nicht zu streng? weil es ist ja eigentlich die Aufgabe von einenm guten Nowcaster diese Abhängigkeit zu modellieren oder?}
Thereby, the distributions are assumed to be independent.
If the nowcasts have the form of a multivariate distribution, including the dependence of the two \acp{pdf}, $f_{t+l|t+l} (x_2)$ has to be replaced by the \ac{pdf} conditional on $x_1$.
As a Monte Carlo approximation of Equation~\eqref{eq:trending-probabilistic-pdf}, the probability can also be calculated by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
For forecasts, the indexes have to be shifted.
If no \acp{pdf} are available, they can be estimated from the \ac{cdf} or quantiles, or the \ac{cdf} or quantiles can be used to generate samples for the Monte Carlo approximation.

>>>>

Third, for measurement, a test method is usually evaluated against a gold standard to test whether the test method could replace the gold standard.
Thus, the gold standard measurements are not available if the quality of the test method is sufficient. \improvement{Änderung des Testgeräts an sich beurteilen}
Thus, the predicted change is computed based on the values of the test method at time $t$ and $t + l$.

@@

\subsection{Accounting for noise and non-informative small changes and bootstrap confidence intervals}\label{subsec:trending-noise}

@@

Confidence intervals can account for the estimation uncertainty of the measures above.
Bootstrap confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
New samples are drawn with replacement from the dataset and are not based on parametric assumptions as classical confidence intervals are.
The confidence interval is then computed based on the bootstrap samples.
We examine three methods for bootstrapping: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
In the \textit{percentile} approach, the confidence interval for the level $\alpha$ is built directly from the empirical distribution of the bootstrap estimators.
The \textit{basic} approach computes the confidence interval based on the non-bootstrap estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but requires smaller datasets for reasonable confidence intervals.
The basic approach balances these two objectives.
We compare the approaches in a small synthetic data study on their small-dataset behavior and computation time in Appendix~\ref{sec:appendix-trending}.
\Ac{bca} is the only method that maintains the confidence level also for small datasets while increasing the computation time only moderately for larger datasets.
Therefore, we use the \ac{bca} method for confidence intervals in the applications in Section~\ref{sec:application}.

@@

\subsection{The conditional trending plot}\label{subsec:trending-cond-prob}

@@

\begin{figure}
    \centering
    \begin{subfigure}{.48\textwidth}
        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_asym_butterfly}
        \caption{First dataset with asymmetric dependence.}
    \end{subfigure}
    \begin{subfigure}{.48\textwidth}
        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_normal}
        \caption{Second dataset. }
    \end{subfigure}
    \caption{Resulting conditional trending plot for different bandwidth selection processes. Cross-validation least squares takes a considerably larger computation time. It does not converge for the first data set and yields a bandwidth too small for the second data set. The rule of thumb is the fastest method but tends to oversmooth. The cross-validation maximum likelihood method yields a more reasonable bandwidth with moderate computation time. \todo{$\varepsilon$ hinzufügen}}\label{fig:trending-cond-prob-bw}
\end{figure}

@@

If the true value is given as a distribution because it is still unknown, the probabilities $p_t$ can be computed by integration.
Let for two nowcasts the distributions be given by \acp{pdf} $f_{t+l|t+l}$ and $f_{t|t+l}$ with \acp{cdf}  $F_{t+l|t+l}$ and $F_{t|t+l}$.
Then, the probability of a negative change can be computed by
\begin{align}
    \bar{p}_t
        &= \int_{\begin{subarray}{l}x_1, x_2 \in \R: \\ x_2 < x_1\end{subarray}} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: (x_1, x_2) \nonumber\\
        &= \int_{x_1 \in \R} \int_{-\infty}^{x_1} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \nonumber\\
        &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \label{eq:trending-probabilistic-pdf}.
\end{align}
\unsure{Die Unabhängigkeitsannahme ist nicht zu streng? weil es ist ja eigentlich die Aufgabe von einenm guten Nowcaster diese Abhängigkeit zu modellieren oder?}
Thereby, the distributions are assumed to be independent.
If the nowcasts have the form of a multivariate distribution, including the dependence of the two \acp{pdf}, $f_{t+l|t+l} (x_2)$ has to be replaced by the \ac{pdf} conditional on $x_1$.
As a Monte Carlo approximation of Equation~\eqref{eq:trending-probabilistic-pdf}, the probability can also be calculated by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
For forecasts, the indexes have to be shifted.
If no \acp{pdf} are available, they can be estimated from the \ac{cdf} or quantiles, or the \ac{cdf} or quantiles can be used to generate samples for the Monte Carlo approximation.
